---
title: "Plots & Models Bleaching MAP Richard"
author: "Richard"
date: "1-6-2021"
output: html_document
---
What do we want to find out?
- Brightness_mean (Bleaching) as an outome variable, how is it influenced by:
Independant Varriables: Species(within?), Experiment (between), Time/Date (within), Structure (between)
- Can we see significant differences per time step, per species, between treatments?
- Is there a correlation with survival?

Libraries --> summary/ varriables --> plots --> Models --> Statistics.
Questions:
1) Plots -->  right information being shown/good enough? 
2) Models --> Is the best mode AIC Lmerand really the best?
3) Models --> What information do we need from the model & how resent it? Table / on plots?
4)any more statistics necesarry?

Import dataset from: 
```{r}
setwd("C:/Users/hp/Documents/R/R-Git-Link/MATbleaching/MATbleaching")
getwd()
```
read data:
```{r}
Bleaching_MAPcsv = read.csv("C:/Users/hp/Documents/R/R-Git-Link/MATbleaching/MATbleaching/Bleaching_MAP.csv", sep = ";")
``` 

```{r}
library(dplyr)
library(tidyverse)
library(rstatix)
library(ggplot2)
library(tidyr)
library(ggpubr)
library(AICcmodavg)
library(lme4)
library(nlme)
library(gratia)
library(mgcv)
```
```{r}
Bleaching_MAP2 <- Bleaching_MAPcsv %>%
rename(
  Brightness = "Green")
```
```{r}
Bleaching_MAP2 <- Bleaching_MAP2 %>%
rename(
  Survival = Survival....)
```
#drop fragment
```{r}
Bleaching_MAP2$Fragment <- NULL
```

```{r}
Bleaching_MAP_Means <- Bleaching_MAP2 %>%
  group_by(Experiment, Structure, Species, Date) %>%
  summarise_all(funs(median, mean), na.rm = TRUE)
```
```{r}
Bleaching_MAP_Means1 <- na.omit(Bleaching_MAP_Means)
```

```{r}
Bleaching_MAP_Means1$Date <- as.Date(Bleaching_MAP_Means1$Date,"%d-%m-%y")
```
#Drop date 10-03-2020:
```{r}
Bleaching_MAP_Means1 <- Bleaching_MAP_Means1 %>%
  filter(Date !="2020-03-10")
```

#CreateID
```{r}
my_cols <- c("Species", "Structure")
Bleaching_MAP_Means1$ID <-do.call(paste, c(Bleaching_MAP_Means1[my_cols], sep = ""))
```

```{r}
Bleaching_MAP_Means1$Experiment <- factor(Bleaching_MAP_Means1$Experiment)
```
```{r}
Bleaching_MAP_Means1$Species <- factor(Bleaching_MAP_Means1$Species)
```
##Check data str
```{r}
str(Bleaching_MAP_Means1)
summary(Bleaching_MAP_Means1)
```
Response to Bleaching per species
```{r}
Fullbxp <- ggplot(Bleaching_MAP_Means1, aes(x= NULL, y=Brightness_mean, group=Species)) +
  geom_boxplot(aes(fill=Species))+
  xlab("Measuring time points(Date)")+
  ylab("Coral Brightness")+
  scale_x_discrete()
Fullbxp
```
Response per speceies per date
```{r fig.width=10}
Fullbxp + facet_grid(.~ Bleaching_MAP_Means1$Date)+
  ggtitle("All Fragments pixel brightness vs time")
```
PLots above plot split per treatment
```{r}
Bleaching_Ctrl <- Bleaching_MAP_Means1 %>%
  filter(Experiment %in% c("CONTROL"))
```
```{r}
Bleaching_MAT <- Bleaching_MAP_Means1 %>%
  filter(Experiment %in% c("MAP"))
```
```{r fig.width=10}
Ctrlbxp <- ggplot(Bleaching_Ctrl, aes(x= NULL, y=Brightness_mean, group=Species)) +
  geom_boxplot(aes(fill=Species))+
  xlab("Measuring time points(Date)")+
  ylab("Coral Brightness")+
  scale_x_discrete()
Ctrlbxp + facet_grid(.~ Date) + ggtitle("Control Fragments pixel brightness vs time") 
```
```{r fig.width=10}
MATbxp <- ggplot(Bleaching_MAT, aes(x= NULL, y=Brightness_mean, group=Species)) +
  geom_boxplot(aes(fill=Species))+
  xlab("Measuring time points(Date)")+
  ylab("Coral Brightness")+
  scale_x_discrete()
MATbxp + facet_grid(.~ Date)+
  ggtitle("MAP Fragments pixel brightness vs time")
```
plots per species
```{r}
Bleaching_Por <- Bleaching_MAP_Means1 %>%
  filter(Species %in% c("Porites cylindrica"))
```
Pocillopora
```{r}
Bleaching_Poc <- Bleaching_MAP_Means1 %>%
  filter(Species %in% c("Pocillopora verrucose"))
```
Acropora
```{r}
Bleaching_Acr <- Bleaching_MAP_Means1 %>%
  filter(Species %in% c("Acropora verweyi"))
```
Millepora
```{r}
Bleaching_Mil <- Bleaching_MAP_Means1 %>%
  filter(Species %in% c("Millepora sp"))
```

```{r fig.width=10}
Acrbxp <- ggplot(Bleaching_MAP_Means1, aes(x= NULL, y=Brightness_mean, group=Experiment)) +
  geom_boxplot(aes(fill=Experiment))+
  xlab("Measuring time points(Date)")+
  ylab("Coral Brightness")+
  scale_x_discrete()
Acrbxp + facet_grid(.~ Bleaching_MAP_Means1$Date)+
  ggtitle("Acropora Fragments pixel brightness vs time")
```
```{r fig.width=10}
Porbxp <- ggplot(Bleaching_Por, aes(x= NULL, y=Brightness_mean, group=Experiment)) +
  geom_boxplot(aes(fill=Experiment))+
  xlab("Measuring time points(Date)")+
  ylab("Coral Brightness")+
  scale_x_discrete()
Porbxp + facet_grid(.~ Bleaching_Por$Date)+
  ggtitle("Porites Fragments pixel brightness vs time")
```
```{r fig.width=10}
Pocbxp <- ggplot(Bleaching_Poc, aes(x= NULL, y=Brightness_mean, group=Experiment)) +
  geom_boxplot(aes(fill=Experiment))+
  xlab("Measuring time points(Date)")+
  ylab("Coral Brightness")+
  scale_x_discrete()
Pocbxp + facet_grid(.~ Bleaching_Poc$Date)+
  ggtitle("Poccilopora Fragments pixel brightness vs time")
```
```{r fig.width=10}
Milbxp <- ggplot(Bleaching_Mil, aes(x= NULL, y=Brightness_mean, group=Experiment)) +
  geom_boxplot(aes(fill=Experiment))+
  xlab("Measuring time points(Date)")+
  ylab("Coral Brightness")+
  scale_x_discrete()
Milbxp + facet_grid(.~ Bleaching_Mil$Date)+
  ggtitle("Milepora Fragments pixel brightness vs time")
```


Models:
```{r}
simplelm <- lm(Brightness_mean~1, Bleaching_MAP_Means1)
simplelm
```
Species as a fixed variable & better model
```{r}
fixedlm <- lm(Brightness_mean~ Species, data=Bleaching_MAP_Means1)
fixedlm
anova(fixedlm)
anova(fixedlm, simplelm)
```
add in a random variable
```{r}
lmerand <- lme(Brightness_mean ~Species*Experiment*Date, data=Bleaching_MAP_Means1, random = ~1|Date)
lmerand
anova(lmerand)
```
```{r}
lmeranDat <- lme(Brightness_mean ~ Species*Experiment , data=Bleaching_MAP_Means1, random = ~1|Date)
lmeranDat
anova(lmeranDat, lmerand)
```
```{r}
lmerand.1 <- lme(Brightness_mean ~1, data=Bleaching_MAP_Means1, random = ~Structure|Date)
lmerand.1
```

```{r}
lme1 <- lme(Brightness_mean ~Species*Experiment*Date, data=Bleaching_MAP_Means1, random = ~ 1|Date/Survival_mean)
lmeDate <- lme(Brightness_mean ~Species*Experiment*Date, data=Bleaching_MAP_Means1, random = ~ Structure|Date)
lmeranSurv <- lme(Brightness_mean ~Species*Experiment*Date, data=Bleaching_MAP_Means1, random = ~ Survival_mean*Structure|Date)
lmeExp <- lme(Brightness_mean ~Species*Experiment*Date, data=Bleaching_MAP_Means1, random= ~1|Date)
anova(lmerand,lme1,lmeranSurv,lmeDate, lmeExp)
```
random=~Survival_mean*Structure|Date better than random=~Structure|Date which is slightly better than not specifying random lmerand

```{r}
lmeranSurv
plot(lmeranSurv)
```

Now to fixed part? -->
Brightness_mean ~ Fixed + Random
```{r}
GLStest <- gls(Brightness_mean~Species*Experiment+Date, data=Bleaching_MAP_Means1, method="REML")
plot(GLStest)
AIC(GLStest)
```
heterogenous but small magnitude


Dealing with heterogeneity/ weighting variables:
```{r}
GLS <- gls(Brightness_mean ~ Species*Date*Experiment, data=Bleaching_MAP_Means1, method = "REML")
op <-par(mfrow =c(1,1),mar=c(4,4,5,5))
plot(GLS,which=c(1),col=1, add.smooth=FALSE,
     caption ="")
plot(Bleaching_MAP_Means1$Date, resid(GLS), xlab= "month", ylab="Residuals")
plot(Bleaching_MAP_Means1$Brightness_mean, resid(GLS), xlab= "Brightness_mean", ylab="Residuals")
par(op)
anova(GLS, GLStest)
```
```{r}
GLSFixed <- varFixed(∼Survival_mean)
M.gls1 <- gls(Brightness_mean ∼ Species*Experiment*Date +(Structure/Date),
weights = GLSFixed, data = Bleaching_MAP_Means1, method ="REML")
anova(GLS, M.gls1)
plot(M.gls1)
```
more residual magnitude but seems more bunched around 0

variance per date (to deal with heterogenity)
```{r}
vf2 <- varIdent(form= ~ 1|Date)
M.gls2 <-gls(Brightness_mean ~ Species*Date*Experiment, data= Bleaching_MAP_Means1, weights = vf2, method = "REML")
anova(GLS, M.gls1, M.gls2)
AIC(GLS2, M.gls1, M.gls2)
plot(M.gls2)
op <-par(mfrow =c(1,1),mar=c(4,4,5,5))
plot(M.gls2,which=c(1),col=1, add.smooth=FALSE,
     caption ="")
plot(Bleaching_MAP_Means1$Date, resid(M.gls2), xlab= "month", ylab="Residuals")
plot(Bleaching_MAP_Means1$Brightness_mean, resid(M.gls2), xlab= "Brightness_mean", ylab="Residuals")
plot(Bleaching_MAP_Means1$Survival_mean, resid(M.gls2), xlab= "Survival_mean", ylab="Residuals")
par(op)

```

```{r}
vf2 <- varIdent(form= ~ Date|Structure)
M.gls3 <-gls(Brightness_mean ~ Species*Date*Experiment, data= Bleaching_MAP_Means1, weights = vf2, method = "REML")
anova(M.gls3, M.gls1, M.gls2)
AIC(M.gls3, M.gls1, M.gls2)
plot(M.gls3)
```
Variation by date than structure?
```{r}
vf3 <- varIdent(form =∼ Structure| Species*Date)
M.gls4 <- gls(Brightness_mean ∼ Date*Experiment*Species,
weights = vf3,data = Bleaching_MAP_Means1, method ="REML")
plot(M.gls4)
anova(M.gls2, M.gls4)
```
```{r}
vf4 <- varIdent(form=~1|Date*Structure)
M.gls6 <- gls(Brightness_mean ∼ Experiment*Species*Date,
weights = vf4,data = Bleaching_MAP_Means1, method ="REML")
plot(M.gls6)
AIC(M.gls2, M.gls4, M.gls6, lmerand3,GLS)
```{r}
M.gam <- gam(Brightness_mean ~ Species*Experiment+s(Date), data=Bleaching_MAP_Means1)
AIC(M.gam, M.gls2)
```
look at if there is grouping by Date: in 
GLS
```{r}
plot(GLS,which = c(1), col = Bleaching_MAP_Means1$Date,
add.smooth = FALSE, caption = "")
```
in MGLS4
```{r}
plot(M.gls4,which = c(1), col = Bleaching_MAP_Means1$Date,
add.smooth = FALSE, caption = "")
```
in lmeranSurv
```{r}
plot(lmeranSurv,which = c(1), col = Bleaching_MAP_Means1$Date,
add.smooth = FALSE, caption = "")
```

```{r}
E <- resid(M.gls4)
coplot(E ∼ Date | Experiment, data = Bleaching_MAP_Means1)
```
```{r}
E <- resid(M.gls1)
coplot( E ~ Date | Experiment, data = Bleaching_MAP_Means1, ylab="Ordinary residuals")
```
```{r}
E <- resid(lmeranSurv)
coplot( E ~ Date | Experiment, data = Bleaching_MAP_Means1)
```
```{r}
E <- resid(lmeranSurv)
coplot( E ~ Date | Species, data = Bleaching_MAP_Means1)
```
```{r}
E <- resid(lmeranSurv)
coplot( E ~ Survival_mean | Experiment, data = Bleaching_MAP_Means1)
```
```{r}
E <- resid(lmeranSurv)
coplot( E ~ Structure | Experiment, data = Bleaching_MAP_Means1)
```
Residuals zien er okee uit denk ik.

Correlation with Survvial nog belangrijk om specifique uit te zoeken?
